Sample Size Determination:\\Posterior Distributions Proximity

### 2024-03-28 paper/main.tex
update

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
\input{preamble}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{color}
\renewcommand\UrlFont{\color{blue}\rmfamily}
\urlstyle{rm}
%
\begin{document}
%
\title{Sample Size Determination:\\Posterior Distributions Proximity}
%
\titlerunning{Posterior Distributions Proximity}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Nikita Kiselev\inst{1} \and Andrey Grabovoy\inst{1}}
\author{Nikita Kiselev \and Andrey Grabovoy}
%
\authorrunning{N. Kiselev, A. Grabovoy}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Moscow Institute of Physics and Technology, Moscow, Russia}
%
\maketitle % typeset the header of the contribution
%
\begin{abstract}
The paper investigates the problem of estimating a sufficient sample size. The issue of determining a sufficient sample size without specifying a statistical hypothesis about the distribution of model parameters is considered. Two approaches to determining a sufficient sample size based on the proximity of posterior distributions of model parameters on similar subsets are suggested. The correctness of the presented approaches is proven in the model with normal posterior distribution. A theorem about moments of the limit posterior distribution of parameters in a linear regression model is proven. A computational experiment is conducted to analyze the properties of the proposed methods.

@@

\begin{theorem}\label{theorem3}
    Let the sets of values of the features and the target variable be bounded, that is, $\exists M\in \mathbb{R}:$ $\|\mathbf{x}\|_2\leqslant M$ and $|y|\leqslant M$. If $\lambda_{\min}\left(\mathbf{X}^{\top}_k \mathbf{X}_k \right) = \omega(\sqrt{k})$ for $k\to \infty$, then in a linear regression model with a normal prior distribution of parameters $\|\mathbf{m}_{k+1} - \mathbf{m}_k\|_2\to 0$ and $\|\mathbf{\Sigma}_{k+1} - \mathbf{\Sigma}_k\|_{F}\to 0$ as $k\to \infty$.
\end{theorem}

@@

Synthetic data is generated from a linear regression model. The number of objects is 500, the number of features is 10. One object is sequentially removed from the given sample until the number of objects in the subsample is equal to the number of features. For each sample size $k$ we calculate the minimum eigenvalue of the matrix $\mathbf{X}_k^\top \mathbf{X}_k$. Also, the values of $KL(k)$ and $S(k)$ are calculated. This process is repeated $B=100$ times.

@@

Fig.~\ref{eigvals} shows the asymptotic behavior of the minimum eigenvalue of the matrix $\mathbf{X}_k^\top \mathbf{X}_k$. We see that when the sample size tends to infinity, the minimum eigenvalue also tends to infinity. Meanwhile, as is necessary for the Theorem~\ref{theorem3}, the graph is higher than $\sqrt{k}$.

@@

@@

\section{Proof of Theorem \ref{theorem1}}
\begin{proof}
The Kullback-Leibler divergence for a pair of normal posterior distributions has the form
    \[ D_{\text{KL}}\left( p_k \| p_{k+1} \right) = \dfrac{1}{2} \left( \mathrm{tr}\left( \mathbf{\Sigma}_{k+1}^{-1} \mathbf{\Sigma}_k \right) + (\mathbf{m}_{k+1} - \mathbf{m}_k)^{\top} \mathbf{\Sigma}_{k+1}^{-1} (\mathbf{m}_{k+1} - \mathbf{m}_k) - n + \log{\left( \dfrac{\det \mathbf{\Sigma}_{k+1}}{\det \mathbf{\Sigma}_{k}} \right)} \right). \]
    Let's express $\mathbf{\Sigma}_{k+1}$ as $\mathbf{\Sigma}_{k+1} = \mathbf{\Sigma}_k + \Delta\mathbf{\Sigma}$. Let's consider each term separately.
    \[ \mathrm{tr}\left( \mathbf{\Sigma}_{k+1}^{-1} \mathbf{\Sigma}_k \right) = \mathrm{tr}\left(\left(\mathbf{\Sigma}_k + \Delta \mathbf{\Sigma} \right)^{-1} \mathbf{\Sigma}_k \right) \to \mathrm{tr}\mathbf{I}_n=n\text{ as } \| \Delta \mathbf{\Sigma} \|_F \to 0, \]
    \[ \left| (\mathbf{m}_{k+1} - \mathbf{m}_k)^{\top}\mathbf{\Sigma}_{k+1}^{-1} (\mathbf{m}_{k+1} - \mathbf{m}_k) \right| \leqslant\| \mathbf{m}_{k+1} -\mathbf{m}_k\|_2^2\|\mathbf{\Sigma}_{k+1}^{-1} \|_2 \ to 0 \text{ as } \| \mathbf{m}_{k+1} - \mathbf{m}_k\|_2 \to 0, \]
    \[ \log{\left( \dfrac{\det \mathbf{\Sigma}_{k+1}}{\det \mathbf{\Sigma}_{k}} \right)} = \log{\left( \dfrac{\det \left( \mathbf{\Sigma}_k + \Delta \mathbf{\Sigma} \right)}{\det\mathbf{\Sigma}_{k}} \right)} \to \log \det\mathbf{I}_n = \log 1 = 0 \text{ as } \| \Delta \mathbf{\Sigma} \|_F\to 0, \]
    from where we have the required.
\end{proof}

\section{Proof of Theorem \ref{theorem2}}
\begin{proof}
Let's use the s-score expression for a pair of normal posterior distributions from \cite{Aduenko2017}:
\[\text{s-score}(p_k, p_{k+1}) = \exp{\left( -\dfrac{1}{2} (\mathbf{m}_{k+1} - \mathbf{m}_k)^{\top} \left( \mathbf{\Sigma}_k + \mathbf{\Sigma}_{k+1} \right)^{-1} (\mathbf{m}_{k+1} - \mathbf{m}_k) \right)}. \]
Because
    \[ \left| (\mathbf{m}_{k+1} - \mathbf{m}_k)^{\top} \left( \mathbf{\Sigma}_k + \mathbf{\Sigma}_{k+1} \right)^{-1} (\mathbf{m}_{k+1} - \mathbf{m}_k) \right| \leqslant \| \mathbf{m}_{k+1} - \mathbf{m}_k \|_2^2 \| \left( \mathbf{\Sigma}_k + \mathbf{\Sigma}_{k+1} \right)^{-1} \|_2 \to 0 \]
    if $\|\mathbf{m}_{k+1} - \mathbf{m}_k\|_2\to 0$, then the value of the quadratic form inside the exponent tends to zero. Therefore, $\text{s-score}(p_k, p_{k+1}) \to 1$ as $\|\mathbf{m}_{k+1} - \mathbf{m}_k\|_2\to 0$.
\end{proof}

@@

    \[ \mathbf{\Sigma} = \left( \alpha \mathbf{I} + \dfrac{1}{\sigma^2} \mathbf{X}^{\top} \mathbf{X} \right)^{-1}, \qquad \mathbf{m} = \left( \mathbf{X}^{\top} \mathbf{X} + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{X}^{\top} \mathbf{y}. \]
Consider the expression $\|\mathbf{\Sigma}_{k+1} - \mathbf{\Sigma}_k\|_2$ norms of difference of covariance matrices for subsamples of size $k$ and $k+1$. Let's introduce the notation $\mathbf{A}_k = \dfrac{1}{\sigma^2}\mathbf{X}^{\top}_k\mathbf{X}_k$. Given the formulas above, we have

@@

    \[ = \sigma^2  \dfrac{1}{\lambda_{\min}\left( \mathbf{X}^{\top}_{k+1} \mathbf{X}_{k+1} \right)} \dfrac{1}{\lambda_{\min}\left( \mathbf{X}^{\top}_k \mathbf{X}_k \right)} \left\| \mathbf{X}^{\top}_{k+1} \mathbf{X}_{k+1} - \mathbf{X}^{\top}_k \mathbf{X}_k \right\|_2. \]

@@

    \[ \left\| \mathbf{X}^{\top}_{k+1} \mathbf{X}_{k+1} - \mathbf{X}^{\top}_k \mathbf{X}_k \right\|_2 = \left\| \sum\limits_{i=1}^{k+1} \mathbf{x}_i \mathbf{x}_i^{\top} - \sum\limits_{i=1}^{k} \mathbf{x}_i \mathbf{x}_i^{\top} \right\|_2 = \left\| \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right\|_2 = \lambda_{\max}\left( \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right) = \]

@@

    \[= \mathbf{x}_{k+1}^{\top} \mathbf{x}_{k+1} = \| \mathbf{x}_{k+1}\|_2^2 \leqslant M^2. \]
By condition $\lambda_{\min}\left(\mathbf{X}^{\top}_k\mathbf{X}_k \right) = \omega(\sqrt{k})$, then $\|\mathbf{\Sigma}_{k+1} - \mathbf{\Sigma}_k\|_2 = o(k^{-1})$ as $k\to \infty$. Next, we will use the equivalence of matrix norms, namely

@@

    \[ \| \mathbf{m}_{k+1} - \mathbf{m}_k \|_2 = \left\| \left( \mathbf{X}_{k+1}^{\top} \mathbf{X}_{k+1} + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{X}_{k+1}^{\top} \mathbf{y}_{k+1} - \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{X}_k^{\top} \mathbf{y}_k \right\|_2 = \]
    Consider that $\mathbf{X}_{k+1}^{\top} = [\mathbf{X}_k^{\top},\mathbf{x}_{k+1}]$ and $\mathbf{y}_{k+1} = [\mathbf{y}_k, y_{k+1}]^{\top}$, then $\mathbf{X}_{k+1}^{\top}\mathbf{X}_{k+1} = \mathbf{X}_k^{\top} \mathbf{X}_k +\mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top}$ and $\mathbf{X}_{k+1}^{\top}\mathbf{y}_{k+1} = \mathbf{X}_k^{\top} \mathbf{y}_k + \mathbf{x}_{k+1} y_{k+1}$.
    \[ = \left\| \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} + \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right)^{-1} \left( \mathbf{X}_k^{\top} \mathbf{y}_k + \mathbf{x}_{k+1} y_{k+1} \right) - \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{X}_k^{\top} \mathbf{y}_k \right\|_2 = \]

@@

\[\left(\mathbf{X}_k^{\top} \mathbf{X}_k+ \alpha\sigma^2 \mathbf{I} +\mathbf{x}_{k+1}\mathbf{x}_{k+1}^{\top}\right)^{-1}= \left(\mathbf{I} + \left( \mathbf{X}_k^{\top}\mathbf{X}_k +\alpha \sigma^2 \mathbf{I}\right)^{-1}\mathbf{x}_{k+1}\mathbf{x}_{k+1}^{\top}\right)^{-1} \left(\mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I}\right)^{-1}.\]

@@

        = \Bigg\| \left[ \left( \mathbf{I} + \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right)^{-1} - \mathbf{I} \right] \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{X}_k^{\top} \mathbf{y}_k + \\ + \left( \mathbf{X}_{k+1}^{\top} \mathbf{X}_{k+1} + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} y_{k+1} \Bigg\|_2 =

@@

        \leqslant \left\| \left( \mathbf{I} + \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right)^{-1} - \mathbf{I} \right\|_2 \left\| \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \right\|_2 \left\| \mathbf{X}_k^{\top} \mathbf{y}_k \right\|_2 + \\ + \left\| \left( \mathbf{X}_{k+1}^{\top} \mathbf{X}_{k+1} + \alpha \sigma^2 \mathbf{I} \right)^{-1} \right\|_2 \left\| \mathbf{x}_{k+1} y_{k+1} \right\|_2

@@

    \[ \left\| \left( \mathbf{I} + \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right)^{-1} - \mathbf{I} \right\|_2 \leqslant \]
    \[ \leqslant \left\| \left( \mathbf{I} + \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right)^{-1} \right\|_2 \cdot \left\| \mathbf{I} \right\|_2 \cdot \left\| \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right\|_2 \leqslant \]

@@

    \[ \leqslant \dfrac{1}{\lambda_{\min}\left( \mathbf{I} + \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right)} \dfrac{\left\| \mathbf{x}_{k+1} \right\|_2^2}{\lambda_{\min}\left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)} \leqslant \]
    \[ \leqslant \dfrac{1}{1 + \lambda_{\min}\left(\left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right)} \dfrac{M^2}{\lambda_{\min}\left( \mathbf{X}_k^{\top} \mathbf{X}_k \right)} \leqslant \]
    The minimum eigenvalue of the product of matrices is estimated by the product of their minimum eigenvalues. In addition, the minimum eigenvalue of the matrix of unit rank $\mathbf{x}_{k+1}\mathbf{x}_{k+1}^{\top}$ is zero.
    \[ \leqslant \dfrac{1}{1 + \lambda_{\max}\left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right) \lambda_{\min}\left( \mathbf{x}_{k+1} \mathbf{x}_{k+1}^{\top} \right)} \dfrac{M^2}{\lambda_{\min}\left( \mathbf{X}_k^{\top} \mathbf{X}_k \right)} = \dfrac{M^2}{\lambda_{\min}\left( \mathbf{X}_k^{\top} \mathbf{X}_k \right)}. \]

@@

    \[ \left\| \left( \mathbf{X}_k^{\top} \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \right\|_2 \left\| \mathbf{X}_k^{\top} \mathbf{y}_k \right\|_2 \leqslant \dfrac{\left\| \mathbf{X}_k^{\top} \mathbf{y}_k \right\|_2}{\lambda_{\min}\left( \mathbf{X}_k^{\top} \mathbf{X}_k \right)} = \dfrac{\left\| \sum\limits_{i=1}^{k} \mathbf{x}_i y_i \right\|_2}{\lambda_{\min}\left( \mathbf{X}_k^{\top} \mathbf{X}_k \right)} \leqslant \dfrac{k M^2}{\lambda_{\min}\left( \mathbf{X}_k^{\top} \mathbf{X}_k \right)} \]

@@

    \[ \left\| \left( \mathbf{X}_{k+1}^{\top} \mathbf{X}_{k+1} + \alpha \sigma^2 \mathbf{I} \right)^{-1} \right\|_2 \left\| \mathbf{x}_{k+1} y_{k+1} \right\|_2 \leqslant \dfrac{M^2}{\lambda_{\min}\left( \mathbf{X}_{k+1}^{\top} \mathbf{X}_{k+1} \right)} \]

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
\input{preamble}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{color}
\renewcommand\UrlFont{\color{blue}\rmfamily}
\urlstyle{rm}
%
\begin{document}
%
\title{Sample Size Determination:\\Posterior Distributions Proximity}
%
\titlerunning{Sample Size Determination: Posterior Distributions Proximity}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Nikita Kiselev\inst{1} \and Andrey Grabovoy\inst{1}}
\author{Nikita Kiselev \and Andrey Grabovoy}
%
\authorrunning{N. Kiselev, A. Grabovoy}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Moscow Institute of Physics and Technology, Moscow, Russia}
%
\maketitle % typeset the header of the contribution
%
\begin{abstract}
The paper investigates the problem of estimating a sufficient sample size. The issue of determining a sufficient sample size without specifying a statistical hypothesis about the distribution of model parameters is considered. Two approaches to determining a sufficient sample size based on the proximity of posterior distributions of model parameters on similar subsets are suggested. The correctness of the presented approaches is proven in the model with normal posterior distribution. A theorem about moments of the limit posterior distribution of parameters in a linear regression model is proven. A computational experiment is conducted to analyze the properties of the proposed methods.

@@

\begin{theorem}\label{theorem3}
    Let the sets of values of the features and the target variable be bounded, that is, $\exists M\in \mathbb{R}:$ $\|\mathbf{x}\|_2\leqslant M$ and $|y|\leqslant M$. If $\lambda_{\min}\left(\mathbf{X}\T_k \mathbf{X}_k \right) = \omega(\sqrt{k})$ for $k\to \infty$, then in a linear regression model with a normal prior distribution of parameters $\|\mathbf{m}_{k+1} - \mathbf{m}_k\|_2\to 0$ and $\|\mathbf{\Sigma}_{k+1} - \mathbf{\Sigma}_k\|_{F}\to 0$ as $k\to \infty$.
\end{theorem}

@@

Synthetic data is generated from a linear regression model. The number of objects is 500, the number of features is 10. One object is sequentially removed from the given sample until the number of objects in the subsample is equal to the number of features. For each sample size $k$ we calculate the minimum eigenvalue of the matrix $\mathbf{X}_k\T \mathbf{X}_k$. Also, the values of $KL(k)$ and $S(k)$ are calculated. This process is repeated $B=100$ times.

@@

Fig.~\ref{eigvals} shows the asymptotic behavior of the minimum eigenvalue of the matrix $\mathbf{X}_k\T \mathbf{X}_k$. We see that when the sample size tends to infinity, the minimum eigenvalue also tends to infinity. Meanwhile, as is necessary for the Theorem~\ref{theorem3}, the graph is higher than $\sqrt{k}$.

@@

\section{Conclusion}

Approaches to determining a sufficient sample size based on the proximity of posterior distributions of model parameters on similar subsamples are proposed. The correctness of the proposed approaches is proved under certain restrictions on the model used. The theorem on the moments of the limit posterior distribution of parameters in a linear regression model is proved. The conducted computational experiment makes it possible to analyze the properties of the proposed methods and their effectiveness.

@@

\section{Proof of Theorem \ref{theorem1}}
\begin{proof}
The Kullback-Leibler divergence for a pair of normal posterior distributions has the form
    \[ D_{\text{KL}}\left( p_k \| p_{k+1} \right) = \dfrac{1}{2} \left( \mathrm{tr}\left( \mathbf{\Sigma}_{k+1}^{-1} \mathbf{\Sigma}_k \right) + (\mathbf{m}_{k+1} - \mathbf{m}_k)\T \mathbf{\Sigma}_{k+1}^{-1} (\mathbf{m}_{k+1} - \mathbf{m}_k) - n + \log{\left( \dfrac{\det \mathbf{\Sigma}_{k+1}}{\det \mathbf{\Sigma}_{k}} \right)} \right). \]
    Let's express $\mathbf{\Sigma}_{k+1}$ as $\mathbf{\Sigma}_{k+1} = \mathbf{\Sigma}_k + \Delta\mathbf{\Sigma}$. Let's consider each term separately.
    \[ \mathrm{tr}\left( \mathbf{\Sigma}_{k+1}^{-1} \mathbf{\Sigma}_k \right) = \mathrm{tr}\left(\left(\mathbf{\Sigma}_k + \Delta \mathbf{\Sigma} \right)^{-1} \mathbf{\Sigma}_k \right) \to \mathrm{tr}\mathbf{I}_n=n\text{ as } \| \Delta \mathbf{\Sigma} \|_F \to 0, \]
    \[ \left| (\mathbf{m}_{k+1} - \mathbf{m}_k)\T\mathbf{\Sigma}_{k+1}^{-1} (\mathbf{m}_{k+1} - \mathbf{m}_k) \right| \leqslant\| \mathbf{m}_{k+1} -\mathbf{m}_k\|_2^2\|\mathbf{\Sigma}_{k+1}^{-1} \|_2 \ to 0 \text{ as } \| \mathbf{m}_{k+1} - \mathbf{m}_k\|_2 \to 0, \]
    \[ \log{\left( \dfrac{\det \mathbf{\Sigma}_{k+1}}{\det \mathbf{\Sigma}_{k}} \right)} = \log{\left( \dfrac{\det \left( \mathbf{\Sigma}_k + \Delta \mathbf{\Sigma} \right)}{\det\mathbf{\Sigma}_{k}} \right)} \to \log \det\mathbf{I}_n = \log 1 = 0 \text{ as } \| \Delta \mathbf{\Sigma} \|_F\to 0, \]
    from where we have the required.
\end{proof}

\section{Proof of Theorem \ref{theorem2}}
\begin{proof}
Let's use the s-score expression for a pair of normal posterior distributions from \cite{Aduenko2017}:
\[\text{s-score}(p_k, p_{k+1}) = \exp{\left( -\dfrac{1}{2} (\mathbf{m}_{k+1} - \mathbf{m}_k)\T \left( \mathbf{\Sigma}_k + \mathbf{\Sigma}_{k+1} \right)^{-1} (\mathbf{m}_{k+1} - \mathbf{m}_k) \right)}. \]
Because
    \[ \left| (\mathbf{m}_{k+1} - \mathbf{m}_k)\T \left( \mathbf{\Sigma}_k + \mathbf{\Sigma}_{k+1} \right)^{-1} (\mathbf{m}_{k+1} - \mathbf{m}_k) \right| \leqslant \| \mathbf{m}_{k+1} - \mathbf{m}_k \|_2^2 \| \left( \mathbf{\Sigma}_k + \mathbf{\Sigma}_{k+1} \right)^{-1} \|_2 \to 0 \]
    if $\|\mathbf{m}_{k+1} - \mathbf{m}_k\|_2\to 0$, then the value of the quadratic form inside the exponent tends to zero. Therefore, $\text{s-score}(p_k, p_{k+1}) \to 1$ as $\|\mathbf{m}_{k+1} - \mathbf{m}_k\|_2\to 0$.
\end{proof}

@@

    \[ \mathbf{\Sigma} = \left( \alpha \mathbf{I} + \dfrac{1}{\sigma^2} \mathbf{X}\T \mathbf{X} \right)^{-1}, \qquad \mathbf{m} = \left( \mathbf{X}\T \mathbf{X} + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{X}\T \mathbf{y}. \]
Consider the expression $\|\mathbf{\Sigma}_{k+1} - \mathbf{\Sigma}_k\|_2$ norms of difference of covariance matrices for subsamples of size $k$ and $k+1$. Let's introduce the notation $\mathbf{A}_k = \dfrac{1}{\sigma^2}\mathbf{X}\T_k\mathbf{X}_k$. Given the formulas above, we have

@@

    \[ = \sigma^2  \dfrac{1}{\lambda_{\min}\left( \mathbf{X}\T_{k+1} \mathbf{X}_{k+1} \right)} \dfrac{1}{\lambda_{\min}\left( \mathbf{X}\T_k \mathbf{X}_k \right)} \left\| \mathbf{X}\T_{k+1} \mathbf{X}_{k+1} - \mathbf{X}\T_k \mathbf{X}_k \right\|_2. \]

@@

    \[ \left\| \mathbf{X}\T_{k+1} \mathbf{X}_{k+1} - \mathbf{X}\T_k \mathbf{X}_k \right\|_2 = \left\| \sum\limits_{i=1}^{k+1} \mathbf{x}_i \mathbf{x}_i\T - \sum\limits_{i=1}^{k} \mathbf{x}_i \mathbf{x}_i\T \right\|_2 = \left\| \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right\|_2 = \lambda_{\max}\left( \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right) = \]

@@

    \[= \mathbf{x}_{k+1}\T \mathbf{x}_{k+1} = \| \mathbf{x}_{k+1}\|_2^2 \leqslant M^2. \]
By condition $\lambda_{\min}\left(\mathbf{X}\T_k\mathbf{X}_k \right) = \omega(\sqrt{k})$, then $\|\mathbf{\Sigma}_{k+1} - \mathbf{\Sigma}_k\|_2 = o(k^{-1})$ as $k\to \infty$. Next, we will use the equivalence of matrix norms, namely

@@

    \[ \| \mathbf{m}_{k+1} - \mathbf{m}_k \|_2 = \left\| \left( \mathbf{X}_{k+1}\T \mathbf{X}_{k+1} + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{X}_{k+1}\T \mathbf{y}_{k+1} - \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{X}_k\T \mathbf{y}_k \right\|_2 = \]
    Consider that $\mathbf{X}_{k+1}\T = [\mathbf{X}_k\T,\mathbf{x}_{k+1}]$ and $\mathbf{y}_{k+1} = [\mathbf{y}_k, y_{k+1}]\T$, then $\mathbf{X}_{k+1}\T\mathbf{X}_{k+1} = \mathbf{X}_k\T \mathbf{X}_k +\mathbf{x}_{k+1} \mathbf{x}_{k+1}\T$ and $\mathbf{X}_{k+1}\T\mathbf{y}_{k+1} = \mathbf{X}_k\T \mathbf{y}_k + \mathbf{x}_{k+1} y_{k+1}$.
    \[ = \left\| \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} + \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right)^{-1} \left( \mathbf{X}_k\T \mathbf{y}_k + \mathbf{x}_{k+1} y_{k+1} \right) - \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{X}_k\T \mathbf{y}_k \right\|_2 = \]

@@

\[\left(\mathbf{X}_k\T \mathbf{X}_k+ \alpha\sigma^2 \mathbf{I} +\mathbf{x}_{k+1}\mathbf{x}_{k+1}\T\right)^{-1}= \left(\mathbf{I} + \left( \mathbf{X}_k\T\mathbf{X}_k +\alpha \sigma^2 \mathbf{I}\right)^{-1}\mathbf{x}_{k+1}\mathbf{x}_{k+1}\T\right)^{-1} \left(\mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I}\right)^{-1}.\]

@@

        = \Bigg\| \left[ \left( \mathbf{I} + \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right)^{-1} - \mathbf{I} \right] \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{X}_k\T \mathbf{y}_k + \\ + \left( \mathbf{X}_{k+1}\T \mathbf{X}_{k+1} + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} y_{k+1} \Bigg\|_2 =

@@

        \leqslant \left\| \left( \mathbf{I} + \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right)^{-1} - \mathbf{I} \right\|_2 \left\| \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \right\|_2 \left\| \mathbf{X}_k\T \mathbf{y}_k \right\|_2 + \\ + \left\| \left( \mathbf{X}_{k+1}\T \mathbf{X}_{k+1} + \alpha \sigma^2 \mathbf{I} \right)^{-1} \right\|_2 \left\| \mathbf{x}_{k+1} y_{k+1} \right\|_2

@@

    \[ \left\| \left( \mathbf{I} + \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right)^{-1} - \mathbf{I} \right\|_2 \leqslant \]
    \[ \leqslant \left\| \left( \mathbf{I} + \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right)^{-1} \right\|_2 \cdot \left\| \mathbf{I} \right\|_2 \cdot \left\| \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right\|_2 \leqslant \]

@@

    \[ \leqslant \dfrac{1}{\lambda_{\min}\left( \mathbf{I} + \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right)} \dfrac{\left\| \mathbf{x}_{k+1} \right\|_2^2}{\lambda_{\min}\left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)} \leqslant \]
    \[ \leqslant \dfrac{1}{1 + \lambda_{\min}\left(\left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right)} \dfrac{M^2}{\lambda_{\min}\left( \mathbf{X}_k\T \mathbf{X}_k \right)} \leqslant \]
    The minimum eigenvalue of the product of matrices is estimated by the product of their minimum eigenvalues. In addition, the minimum eigenvalue of the matrix of unit rank $\mathbf{x}_{k+1}\mathbf{x}_{k+1}\T$ is zero.
    \[ \leqslant \dfrac{1}{1 + \lambda_{\max}\left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right) \lambda_{\min}\left( \mathbf{x}_{k+1} \mathbf{x}_{k+1}\T \right)} \dfrac{M^2}{\lambda_{\min}\left( \mathbf{X}_k\T \mathbf{X}_k \right)} = \dfrac{M^2}{\lambda_{\min}\left( \mathbf{X}_k\T \mathbf{X}_k \right)}. \]

@@

    \[ \left\| \left( \mathbf{X}_k\T \mathbf{X}_k + \alpha \sigma^2 \mathbf{I} \right)^{-1} \right\|_2 \left\| \mathbf{X}_k\T \mathbf{y}_k \right\|_2 \leqslant \dfrac{\left\| \mathbf{X}_k\T \mathbf{y}_k \right\|_2}{\lambda_{\min}\left( \mathbf{X}_k\T \mathbf{X}_k \right)} = \dfrac{\left\| \sum\limits_{i=1}^{k} \mathbf{x}_i y_i \right\|_2}{\lambda_{\min}\left( \mathbf{X}_k\T \mathbf{X}_k \right)} \leqslant \dfrac{k M^2}{\lambda_{\min}\left( \mathbf{X}_k\T \mathbf{X}_k \right)} \]

@@

    \[ \left\| \left( \mathbf{X}_{k+1}\T \mathbf{X}_{k+1} + \alpha \sigma^2 \mathbf{I} \right)^{-1} \right\|_2 \left\| \mathbf{x}_{k+1} y_{k+1} \right\|_2 \leqslant \dfrac{M^2}{\lambda_{\min}\left( \mathbf{X}_{k+1}\T \mathbf{X}_{k+1} \right)} \]
